---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

### Loading the libraries

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
```


## Loading and preprocessing the data
```{r, echo = TRUE}
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
```


## Sum of steps per day
```{r, echo = TRUE}
sum <- df %>% filter(!is.na(steps)) %>% group_by(date) %>% summarize(sum=sum(steps))
sum
```

### Histogram of the total number of steps taken each day
```{r, echo = TRUE}
ggplot(sum, aes(x=sum))+geom_histogram(fill="red",binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

### Mean of the total number of steps taken per day
```{r, echo = TRUE}
mean_steps <- mean(sum$sum, na.rm=T)
mean_steps
```
### Median of the total number of steps taken per day
```{r, echo = TRUE}
median_steps <- median(sum$sum, na.rm=T)
median_steps
```

### Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r, echo = TRUE}
actpat <- df %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(mean=mean(steps))
ggplot(actpat,aes(x=interval,y=mean))+ geom_line(colour="blue")

```
 
### 5-minute interval which on average across all the days in the dataset contains the maximum number of steps
```{r, echo = TRUE}
actpat[which.max(actpat$mean),]
```

### Number of missing values
```{r, echo = TRUE}
sum(is.na(df))
```
### Number of missing values per column
```{r, echo = TRUE}
sapply(df, function(x) sum(is.na(x)))
```

### Replacing NA's with the mean of the 5-minute interval
```{r, echo = TRUE}
data <- df
nas <- is.na(df$steps)
avg_interval <- tapply(data$steps, data$interval, mean, na.rm=TRUE, simplify=TRUE)
data$steps[nas] <- avg_interval[as.character(data$interval[nas])]
```

### Sum of total steps per day after removing NAs.
```{r, echo = TRUE}
sum2 <- data %>% filter(!is.na(steps)) %>% group_by(date) %>% summarize(sum=sum(steps))
sum2
```

### Histogram of dataset without NA.
```{r}
ggplot(sum2, aes(x=sum))+geom_histogram(fill="red",binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")

```

### Mean after removing NAs
```{r, echo = TRUE}
mean_steps2 <- mean(sum2$sum, na.rm=T)
mean_steps2
```
### Median after removing NAs
```{r, echo = TRUE}
median_steps2 <- median(sum2$sum, na.rm=T)
median_steps2
```

### Table of mean and median before and after removing NAs.
```{r, echo = TRUE}
diff <- data.frame(c(mean_steps, mean_steps2),c(median_steps,median_steps2))
colnames(diff) <- c("mean","median")
rownames(diff) <- c("with NA", "without NA")
diff
```

### Creating a new column with the type of day regarding is it weekday or weekend
```{r, echo = TRUE}
data <- data %>% mutate(weektype=ifelse(weekdays(date)=="Saturday" | weekdays(date)=="Sunday","Weekend","Weekday"))
```

### Plot of  the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r, echo = TRUE}
data2 <- data %>% group_by(interval, weektype) %>% summarise(mean=mean(steps))
ggplot(data2,aes(x=interval,y=mean))+ geom_line(aes(colour=weektype)) + facet_wrap(weektype~.,ncol = 1, nrow=2)
```


